---
title: "R_programming_w2"
output: html_document
---

#R programming
## week 2
###R concole input and evaluation

```{r,echo=TRUE}
x<-1
print(x)
x ##print x directly,auto-printing
msg<-"hello"
print(msg)
y<-1:20
```



### data types
####object
- character
- numeric(real number)
- integer
- complex
- logical(true/false)

one class in one vector except **list**

####number
- 1: a numeric object; 1L:an integer
- inf: infinity   1/0=inf
- nan: no a number 0/0=nan

####attributes
- names, dimnames
- dimensions
- class
- class
- length

####vector
```{r,echo=FALSE}
a<-c(0.5,0.6)    ##numeric
b<-c(TRUE,FALSE) ##logic
c<-c(T,F)        ##logic
d<-c("a","b","c")##character
e<-9:29          ##integer
f<-c(1+0i,2+4i)  ##complex

a
b
c
d
e
f
g<-vector("numeric",length = 10)
g
```

```{r,echo=FALSE}
a<-c(1.7,"a")
b<-c(TRUE,2)
c<-c("a",TRUE)
a
b
c
```

```{r}
##explicit coercion
x<-0:5 
print(class(x))
as.numeric(x)
as.logical(x)
as.character(x)
```

####list
```{r}
x<-list(1,"a",TRUE,1+4i)
x ##double braket
```

####matrices
```{r}
m<-matrix(nrow=2,ncol=3)
m
dim(m)
attributes(m)

n<-matrix(1:6,nrow=2,ncol=3)
n

p<-1:10
p
dim(p)<-c(2,5)
p
```

```{r}
x<-1:3
y<-10:12
cbind(x,y)
rbind(x,y)
```


####factors
```{r}
x<-factor(c("yes","yes","no","yes","no"))
x
table(x)
unclass(x)
attr(x,"levels") ## wrong in PPT
##because "no" is before "yes" in alphbet, so "no" is 1 and "yes" is 2

y<-factor(c("yes","yes","no","yes","no"),
          levels = c("yes","no"))
y
##set the factor "yes"=1, "no"=2
unclass(y)
attr(y,"levels")
```

#### missing values
- is.na()
- is.nan()
- NA value have class, like integer NA, character NA, etc  NA = not available
- a NaN value is also NA but the converse is not true  NaN = not a number
```{r}
x<-c(1,2,NA,10,3)
is.na(x)
is.nan(x)
y<-c(1,2,NaN,NA, 3)
is.na(y)
is.nan(y)
```

#### data frames
- same length (!!)
- different classes of objects in each column(unlike matrix)
- row.names
- read.table()/read.csv()
- can be converted to a matrix: data.matrix()
```{r}
x<-data.frame(foo=1:4,bar=c(T,T,F,F)) 
   ## 2 columes (foo, bar), and four records (rows)
x
nrow(x)
ncol(x)
```

#### the name attribute
```{r}
x<-1:3
names(x)
names(x)<-c("foo","bar","norf")
x
y<-list(a=1,b=2,c=3)
y

m<-matrix(1:4,nrow=2,ncol=2)
dimnames(m)<-list(c("a","b"),c("c","d")) ## firstly apply to rows, then columns
m
```

### Reading Tabular Data
1. reading data
- read.table, read.csv
- readLines,
- source
- dget
- load
2. writing data
- write.table
- writeLines

### Reading large dataset
- R will run faster if it's told what each class is 
- read some part of the data; get class for each colume; read the whole data using the class

innitial <- read.table("xx.txt".nrows=100)
classes <- sapply(initial,classes)
tabAll <- read.table("xx.txt",colClasses=classes)

### texture data format
- dput-ting R Objects(write R code)
```{r}
y<-data.frame(a=1,b="a")
dput(y)
dput(y,file='y.R')
new.y<-dget("y.R")
new.y
```

- dumping R objects
```{r}
x<-"foo"
y<-data.frame(a=1,b="a")
dump(c("x","y"),file="data.R")
rm(x,y)
source("data.R")
y
x

```

###connections: interfaces to the outside world
connections
-file
-gzfile: open compress files
-bzfile
-url

con<-file("xxx.txt","r")
data<-read.csv(con)
close(con)

data<-read.csv("foo.txt")

readling lines of a text file

con<-gzfile("words.gz")
x<-readlines(con,10) #only read top 10 lines

con<-url("http://....")
x<-readLines(con)
head(x)


### subsetting R objects: basic
```{r}
x<-c("a","b","c","c","d","a")
x[1]
x[2]
x[1:4]
x[x>"a"] #logic
u<-x>"a"
x[u]
```

### subsetting R objects: list
```{r}
x<-list(foo=1:4,bar=0.6,baz="hello")
x[1] # a list that contains 1-4
x[[1]] # just the sequence
x$bar
x[["bar"]] #just the content
x["bar"] # a list

name<-"foo"
x[[name]]
```

```{r}
x<-list(a=list(10,12,14),b=c(3.14,2.81))
x[[c(1,3)]]
x[[1]][[3]]
x[c(2,1)]
x[[2]][[1]]
```

### subsetting R objects: matrix
```{r}
x<-matrix(1:6,2,3)
x[1,2]
x[2,1]
x[1,] #just get numbers
x[,2]
X[1,2,drop=FALSE] #get a matrix
x[1,2,drop=TRUE] #it's default
x[1,,drop=FALSE] #get a matrix
```

###ubsetting R objects: partial matching
```{r}
x<-list(aardvark=1:5)
x$a #partical matching -> aardvark
x[["a"]] #no result,[[]]must be exact by default
x[["a",exact=FALSE]] #partical matiching -> aardvark
```

###subsetting R objects: removing NA values
```{r}
x<-c(1,2,NA,4,NA,5)
bad<-is.na(x)
x[!bad]
y<-c("a","b",NA,"d",NA,"f")
good<-complete.cases(x,y)
z<-c("h","i","j","k",NA,NA)
good2<-complete.cases(x,y,z)
```


### vectorized operations
```{r}
x<-1:4;y<-6:9
x+y # calculate paralle
x>2 # compare every one
x==2
y==8
x/y
```

```{r}
x<-matrix(1:4,2,2)
y<-matrix(rep(10,4),2,2) #four 10, 2*2

x*y #element-wise multiplication
x/y 
x%*%y #true matrix multiplication
```
